---
format: gfm
title: Benchmarks
---

A basic benchmark of `opentripplanner` and `rvalhalla` is shown below.

```{r}
#| eval: true
#| message: false
knitr::kable(readr::read_csv("bench.csv"))
```

Note: this benchmark is not comparing like with like, and is designed to be a basis for future benchmarks, perhaps with other routing engines and other interfaces with reference to more realistic use cases.
It does show that `rvalhalla` is reasonably fast for the simple case of routing between two points.

See the reproducible code below to generate the benchmark at home.

# Setup


```{r}
remotes::install_cran("opentripplanner")
library(opentripplanner)
library(sf)
knitr::opts_chunk$set(eval = FALSE)
```


# opentripplanner

OTP relies on Java 17, which can be installed on Ubuntu as follows:

```bash
sudo apt install openjdk-17-jdk
# Set the version of Java to use:
sudo update-alternatives --config java
```


```{r}
library(opentripplanner)
# Path to a folder containing the OTP.jar file, change to where you saved the file.
path_data = file.path(tempdir(), "OTP")
dir.create(path_data)
path_otp = otp_dl_jar(version = "2.2.0")
otp_dl_demo(path_data)
# Build Graph and start OTP
log1 = otp_build_graph(otp = path_otp, dir = path_data)
log2 = otp_setup(otp = path_otp, dir = path_data)
otpcon = otp_connect(timezone = "Europe/London")
```

Following the routing tutorial in [github.com/itsleeds/TDS](https://github.com/ITSLeeds/TDS/blob/ff0c7346d2f872539faae11224aaa76b79e8c2b6/practicals/6-routing.Rmd), let's generate some routes:


```{r, message=FALSE}
u = "https://github.com/ITSLeeds/TDS/releases/download/22/NTEM_flow.geojson"
desire_lines = read_sf(u)
desire_lines = desire_lines[1:9, ]
fromPlace = lwgeom::st_startpoint(desire_lines)
toPlace = lwgeom::st_endpoint(desire_lines)
fromPlace = st_sf(data.frame(id = desire_lines$from, geometry = fromPlace))
toPlace = st_sf(data.frame(id = desire_lines$to, geometry = toPlace))
```

Calculate a route as follows:

```{r, message=FALSE,  eval=FALSE}
route = otp_plan(otpcon, 
                  fromPlace = c(-1.17502, 50.64590), 
                  toPlace = c(-1.15339, 50.72266))
```

Calculate all routes as follows:

```{r, message=FALSE}
#| echo: false
#| eval: false
fromPlace = sf::st_coordinates(fromPlace)
toPlace = sf::st_coordinates(toPlace)
routes_drive_top = otp_plan(otpcon = otpcon,
                            fromPlace = fromPlace[1:9, ],
                            toPlace = toPlace[1:9, ])
```

We can do the same with rvalhalla as follows:

```{r, message=FALSE}
library(osmextract)
iow_url = oe_match("Isle of Wight")
iow_url
dir.create("custom_files_iow", showWarnings = FALSE)
oe_download(
  file_url = iow_url$url,
  file_size = iow_url$file_size,
  download_directory = "custom_files_iow/",
  file_basename = "isle_of_wight"
)
```

``` bash
# Stop the running container:
docker stop valhalla_gis-ops
# Remove the container:
docker rm valhalla_gis-ops
docker rmi ghcr.io/gis-ops/docker-valhalla/valhalla:latest
docker run -dt --name valhalla_gis-ops -p 8002:8002 -v $PWD/custom_files:/custom_files -e tile_urls=https://download.geofabrik.de/europe/great-britain/england/isle-of-wight-latest.osm.pbf ghcr.io/gis-ops/docker-valhalla/valhalla:latest
```

```{r}
#| echo: false
#| eval: false
route1 = vh_route(andorra_la_vella, pas_de_la_casa, costing = "pedestrian")
```

```{r, message=FALSE}
devtools::load_all()
library(rvalhalla)
route_vh = vh_route(c(-1.17502, 50.64590), c(-1.15339, 50.72266), costing = "pedestrian")
plot(route_vh)
```


```{r, message=FALSE}
bench = bench::mark(check = FALSE,
  otp = otp_plan(otpcon, 
                  fromPlace = c(-1.17502, 50.64590), 
                  toPlace = c(-1.15339, 50.72266)),
  rvalhalla = vh_route(c(-1.17502, 50.64590), c(-1.15339, 50.72266), costing = "pedestrian")
)
readr::write_csv(bench[c(1, 4, 5)], "bench.csv")
```
